Verifying Hierarchical Cache Coherence Protocols: a theorem-prover based modular, reusable proof

### 2014-02-04 paper/Deadlock.tex

In Section \ref{safety} we showed how the MSI protocol of Figure \ref{trans}
obeys the Store atomicity theorem. In addition to this criteria, we must
establish that the system makes forward progress.

Forward progress can be either of the following:
\begin{enumerate}
\item Weak forward progress: The system never reaches a state where none of the transitions of Figure
\ref{trans} apply or
\item Strong forward progress: At any time, eventually a new request from some processor will be
processed
\end{enumerate}

In order to prove strong forward progress, we need to
establish some more conditions on the strategy to execute the transitions of
Figure \ref{trans} in each cache -- the caches must ensure local
starvation-freedom in the processing of messages when sufficient resources are
available. We do not discuss this issue in this paper and instead focus on 
weak forward progress.

>>>>

In Section~\ref{safety} we showed how the MSI protocol of
Figure~\ref{trans} obeys the store atomicity theorem, showing that
only correct behavior occurs. To complete our correctness picture, we
must also show that we avoid deadlock, \ie\ we make forward progress.

Forward progress can be either of the following:
\begin{enumerate}
\item Weak forward progress: The system never reaches a state where
  none of the transitions of Figure~\ref{trans} apply or
\item Strong forward progress: At any time, eventually a new request
  from some processor will be processed
\end{enumerate}

To prove strong forward progress given weak fairness, we need some
guarantees on the choice of executions of the transitions
Figure~\ref{trans} in each cache. Specifically, caches must ensure
local starvation-freedom in the processing of messages when sufficient
resources are available. This can be enforced via during hardware
compilation; for this paper we will focus only on weak forward
progress.

### 2014-02-04 paper/Introduction.tex
intuition

Because verified protocols are often described via Term Rewrite
Systems (TRS). It was shown by Dave
\etal{}~\cite{DNA:CoherenceImplementation} that such a verified
protocol could be transliterated into Bluespec, a hardware description
language based on Guarded Atomic Actions, a restriction of
TRS~\cite{Hoe:TCAD}. This reduces the correctness of the
implementation to the correctness of the protocol, effectively
removing any implementation bugs. While the current Bluespec compiler
has certain restrictions which limit the expressiveness of certain
operations \eg{} fairness is not guaranteed, atomic actions are always
implemented in a single cycle, work has been done to remove these
restrictions~\cite{DNA:CoherenceImplementation, Karczmarek}.

@@

\noindent\textbf{Paper Outline}. In Section~\ref{Sec:Background} we
discuss our notion of correctness for our cache coherence protocol
framework. We discuss our MSI protocol example and its implementation
in Section~\ref{Sec:Protocol}. In Section~\ref{Sec:ProofStructure} we discuss
the structuring of our proof framework and verification of the base
protocol. We show implementation refinements in
Section~\ref{Sec:Refinements} and conclude with a general discussion
of the applicability of our technique.

>>>>

Because verified protocols are often described via Term Rewrite
Systems (TRS). It was shown by Dave
\etal{}~\cite{DNA:CoherenceImplementation} that such a verified
protocol could be transliterated into Bluespec~\cite{Bluespec:TFRG}, a
hardware description language based on Guarded Atomic Actions, a
restriction of TRS~\cite{Hoe:TCAD}. This reduces the correctness of
the implementation to the correctness of the protocol, effectively
removing any implementation bugs. While the current Bluespec compiler
has certain restrictions which limit the expressiveness of certain
operations \eg{} fairness is not guaranteed, atomic actions are always
implemented in a single cycle, work has been done to remove these
restrictions~\cite{DNA:CoherenceImplementation, Karczmarek}.

@@

\noindent\textbf{Paper Outline}. In Section~\ref{Sec:Background} we
discuss our notion of correctness for our cache coherence protocol
framework. We discuss our MSI protocol example and its implementation
in Section~\ref{Sec:System}. In Section~\ref{Sec:ProofStructure} we discuss
the structuring of our proof framework and verification of the base
protocol. We show implementation refinements in
Section~\ref{Sec:Refinements} and conclude with a general discussion
of the applicability of our technique.
